Sampling Strategies in Decisions from Experience

### 2022-02-14 manuscript.Rmd
Updates CPT

The human mind is a cognitive system that operates on inputs from an environment or memory.
In the domain of decision making, such inputs provide information about the choice alternatives among which people can choose, hereafter treated as elements $c$ of the set $C$.
That is, each choice alternative $c \in C$ is assigned some information from the inputs and this information may be used by the mind to learn about the *properties* of the respective alternative.
In turn, decision making can be considered an information processing problem which starts from the set $C$, containing all choice alternatives, and stops at a proper subset $C' \subset C$ with the inclusion map
$$
\begin{aligned}
  \iota : C' &\to C \\
  c &\mapsto c  \quad \forall c \in C'
\end{aligned}
\tag{1}
$$
[@jostMathematicalConcepts2015, p. 15], containing only the chosen alternatives.
Indeed, irrespective of their exact mathematical specification or computational implementation, decision models generally take as inputs information about a set of choice alternatives to return a subset thereof as outputs [@heOntologyDecisionModels2020].

@@

# The Description of Choices in Cumulative Prospect Theory

Since there is an uncountable number of decisions people confront in the natural world, behavioral decision research routinely abstracts from decisions between particular choice alternatives, e.g., the choice between job offers, political parties to vote for, investment plans, and whatnot.
Rather, with the choice between at least two *prospects* [see @kahnemanProspectTheoryAnalysis1979, see also below], it studies a case which is counterfactual in that it omits the many particularities of each decision, but retains in the form of prospects the fundamental properties that almost all choice alternatives are assumed to possess [@lopesThoughtsPsychologicalConcept1983].^[See @knightRiskUncertaintyProfit1921, Chapter 1, for a timeless discussion of the reasonableness of such abstractions.] 
These fundamental properties are considered to be the possible outcomes of a choice alternative---hereafter denoted by $x$---and the probabilities with which these outcomes occur following the choice of the alternative---hereafter $p_X(x)$.
Roughly speaking, then, all choice alternatives that can be fully described by their outcomes and probabilities are prospects (or: gambles), which are exclusively considered below. 
<!--For illustration, take the choice alternatives in the following example with monetary outcomes:

@@

Now, let the exact axiomatic definition of a prospect be that of a discrete random variable's distribution.
Let therefore the tuple $(\Omega, \mathcal{F}, P)$ be a probability space. 
Following the @kolmogorovFoundationsTheoryProbability1950 axioms, $\Omega$ is the sample space containing a finite set of elements
$$
\omega_i \in \{\omega_1, ..., \omega_n\} = \Omega
\; .
\tag{2}
$$ 
$\mathcal{F}$ is the set of all possible subsets of $\Omega$, i.e., the power set $\mathcal{P}(\Omega)$ with each element $A \in P(\Omega)$ being a subset of $\Omega$.^[Note that this paper considers only the case of finite sample spaces. The exact axiomatic definition of $\mathcal{F}$, which can be found in Kolmogorov [-@kolmogorovFoundationsTheoryProbability1950, pp. 2-3 and 14-15], also covers infinite sample spaces and is that of a $\sigma$-algebra. For the finite case, however, $\mathcal{P}(\Omega)$ can be considered as such [@georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1.1].] 
$P$ is a probability measure
$$
P: \mathcal{F} \to [0,1]
\; ,
\tag{3}
$$
which assigns each subset $A \in \mathcal{F}$ a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.<!--^[The probabilities $p(\omega)$ are determined by the probability mass function $p : \Omega \to [0,1]$, which is used to construct the probability measures $P(A) = \sum_{\omega \in A} p(\omega)$ for all $A \in \mathcal{F}$.]-->
The random variable $X$ is then a measurable function
$$
\begin{aligned}
  X : (\Omega, \mathcal{F}) &\to (\Omega', \mathcal{F'})  \\
  \omega &\mapsto x
  \; ,
\end{aligned}
\tag{4}
$$
where $(\Omega', \mathcal{F'})$ is a measurable space with each element $x \in \Omega'$ being a possible outcome of a prospect and each element $A' \in \mathcal{F'}$ being a subset of $\Omega'$.
An outcome's probability $p_X(x)$ is then provided by the probability measure
$$
P_X : \mathcal{F'} \to [0,1]
\tag{5}
$$
with 
$$
P_X(A' \in \mathcal{F'}) := P(X^{-1} \times A') = P(\{\omega: X(\omega) = x \in A'\})
\tag{6}
$$
[see @kolmogorovFoundationsTheoryProbability1950, p. 21; see also @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1, for a comprehensive introduction to random variables].
That is, the set consisting of a single outcome $x$ can be treated as a subset $\{x\} \in \mathcal{F'}$ and one therefore obtains $p_X(x) = P(X(\omega) = x)$.
Note that the probability measure in (5) and (6), respectively, is the discrete random variable's distribution and therefore a prospect.
To summarize and simplify, given a choice between $1, ..., k$ prospects, the inputs from the environment or memory should provide information about the set
$$
C = \{P_{X_1}, ... , P_{X_k}\}
\tag{7}
$$
with each element $c \in C$ being a distribution $P_X$ of a random variable defined on a different probability space.
Accordingly, each $c$ can itself be considered a set of outcome-probability pairs $(x_i, p_X(x_i)) \in \{(x_1, p_X(x_1)), ..., (x_n, p_X(x_n))\}$.

However, rarely in peoples' daily life, inputs take the form of explicit descriptions of all outcomes-probability pairs<!--(as in Example 1)-->.
In such cases, people would make *decisions from description* (DfD), since the information from which the mind learns about the prospects are complete descriptions thereof.
Yet, rather often, people make *decisions from experience* [DfE, @hertwigDecisionsExperienceEffect2004], where the mind can learn about the prospects only from a record of past experiences with them<!--, which necessarily provides only incomplete and probable information-->.
That is, people may have experienced the possible outcomes of a prospect when they themselves or others have made the same decision repeatedly in the past, and the record of experienced outcomes for this particular prospect is a rough description of a *sample*. 
<!--The mind may then infer from the relative frequency with which an outcome occurs in such a sample of past decisions its probability of occurrence for a future decision. 
Of course, ceteris paribus, the accuracy of this inference depends on the size of the sample.
For illustration, take the sampled outcomes in the following example, where alternative A2 from Example 1 was chosen ten times without prior knowledge about the possible outcomes and their actual probabilities:

\noindent
**Example 2.**
```{r echo=TRUE}
set.seed(1)
sample(a2_outcomes, size = 10, prob = a2_prob, replace = TRUE)
```

\noindent
One may now make an inference from the sample about the possible outcomes of alternative A2 and their probabilities. 
However, this inference is only probable, since the relative frequencies of outcomes will almost surely change with increasing sample size.

-->
More formally, let the sample of a prospect be a set of realizations of the respective random variable $X$ according to its distribution $P_X$.
Accordingly, for the choice between $1, ..., k$ prospects, the inputs to DfE are generated by a set $S$ of stochastic processes
$$
S = \{X_{1t}, ... , X_{kt}\}
\tag{8}
$$
with $t \in \{1, ..., N\}$ denoting a sequence of independent and identically distributed (i.i.d.) random variables and $N$ denoting the size of the respective sequence.^[Note that this paper considers the counterfactual case where samples are random, although it has been emphasized that in the natural world samples are "virtually never random" [@fiedlerBewareSamplesCognitiveecological2000, p. 660]. To account for non-random sampling, e.g., in the form of sampling biases [e.g., @fiedlerBewareSamplesCognitiveecological2000] or dynamic information structures [e.g., @cohenEffectPerceivedPatterns2021; @plonskyRelianceSmallSamples2015], the i.i.d-assumption would need to be dropped, which is out of the scope of this paper.]
The sets of realizations that result from these stochastic processes may then be used to make an uncertain inference about the respective sets of outcome-probability pairs in (7).

The choices people actually make between prospects---be that in DfD or DfE---are often described in terms of deviations from expected value (EV) maximization, where the EV of a prospect is given by 
$$
EV = \sum_i^n x_i \times p_X(x_i)
\; .
\tag{9}
$$ 
That is, following the principle of EV maximization, one would first calculate the prospects' EV according to (9) and thereafter chose the prospect with the largest EV.
To *describe* how peoples' choices deviate from EV maximization [see @erevAnomaliesForecastsDescriptive2017, for a recent replication of classical demonstrations], the models of cumulative prospect theory [CPT, @tverskyAdvancesProspectTheory1992] and similar rank-dependent models [see @stottCumulativeProspectTheory2006, for an overview of models] fit choice data with a combination of two parameterized, monotonic functions that may replace the objective outcomes and probabilities in (9) by nonlinear transformations thereof. 
More specifically, the models of CPT fit choice data by maximizing the value 
$$
V = \sum_i^n v(x_i) \times \pi_i
\; ,
\tag{10}
$$ 
with objective outcomes $x_i$ being transformed by a *value function* $v$, and cumulative decision weights $\pi_i$ being derived from cumulative probabilities of the distribution $P_X$ that may be transformed by a *weighting function* $w$ (see (11) and (12) below).^[Note that in CPT, the value and weighting functions are sign-dependent. Accordingly, two value and weighting functions are needed for prospects with positive and negative outcomes, respectively [see @tverskyAdvancesProspectTheory1992]. Note further that CPT or other originally deterministic decision models may be amended by a stochastic choice function to account for the probabilistic nature of choices [see, e.g., @rieskampProbabilisticNaturePreferential2008].]

## Probability Weighting and The Description-Experience Gap

According to @tverskyAdvancesProspectTheory1992, let the cumulative decision weights be defined by
$$
\pi_i =
  \begin{cases}
     w^+(P(X \geq x_i)) - w^+(P(X > x_i)) \quad \forall x \geq 0, \\
     w^-(P(X \leq x_i)) - w^-(P(X < x_i)) \quad \text{else}, 
  \end{cases}
\tag{11}
$$
with $w^+(0) = w^-(0) = 0$ and $w^+(1) = w^-(1) = 1$.
Now, given that $w(P_X(\Omega')) = w(1) = 1$, it follows from (11) that for any prospect that is strictly positive or strictly negative, the weighting function may redistribute the entire probability mass $P_X(\Omega') = 1$ across outcomes [see @zilkerNonlinearProbabilityWeighting2021]. 
As a consequence, the decision weights may deviate---i.e., they may be smaller or greater---from the respective objective probabilities<!-- and these deviations are usually regressive-->.
Although the possible deviation patterns, which span the entire probability scale, depend on the exact parameterization of the weighting function [e.g., @goldsteinExpressionTheoryPreference1987; @tverskyAdvancesProspectTheory1992; @prelecProbabilityWeightingFunction1998; see also @stottCumulativeProspectTheory2006], they allow the adoption of an as-if weighting terminology for the description of choice behavior, which was originally introduced outside of the CPT models [see @pachurWeightUncertainEvents2019].
Figure 1 illustrates such a description in CPT with different parameter settings for the two-parameter probability weighting function of @goldsteinExpressionTheoryPreference1987 with
$$
\begin{aligned}
  w : [0,1] &\to [0,1] \\
  p_X(x) &\mapsto \frac{\delta \times p_X(x)^{\gamma}}
  {\delta \times p_X(x)^{\gamma} + (1-p_X(x))^{\gamma}}
  \; ,
\end{aligned}
\tag{12}
$$
where $\gamma \in [0,2]$ and $\delta \geq 0$ determine the function's curvature and elevation, respectively.
Essentially, if the estimate of a decision weight is larger (smaller) than the objective probability from which it is derived, the weighting function runs above (below) the dashed identity line.
The choice data are then described *as if* people maximized the value in (8) and applied the over- and underweighting pattern indicated by the estimated shape of the overall probability weighting function.
Importantly, the as-if prefix indicates that it cannot be concluded from the inferred decision weights that the mind does perform a probability weighting computation in its literal sense, but it processes the inputs in a way that the outputs approximate the solutions of what can be considered probability weighting on an abstract computational level [see, e.g., @griffithsRationalUseCognitive2015].

@@

ggplot(wf) +
  geom_line(aes(prob, w, group = gamma, color = gamma)) +
  geom_segment(x = 0, xend = 1, y =  0, yend = 1, linetype = "dashed", size = .3) +
  scale_color_scico(palette = "cork") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x = expression(paste("p(", omega, ")")),
       y = expression(paste("w(p(", omega, "))")), 
       color = expression(gamma)) + 
  facet_wrap(~delta, scales = "fixed") + 
  theme_apa() 
```

@@

@@

The argument so far has been quite general and abstract, since there is no complete answer to the question raised above this paragraph---there is, of course, an uncountable number of decisions and so are there choice alternatives and properties.
Therefore, in the present paper, *prospects* are used as a proxy for the variety of choice alternatives the mind is required to evaluate in natural world decisions. 
Consequently, the argument continues to proceed at an abstract level: with choices between at least two prospects, the current paper considers a case which is counterfactual, since it is freed from many varieties and complexities that the mind is normally confronted with. 
The case is, however, quite paradigmatic in research on decision theory and there are good reasons for that [cf. @knightRiskUncertaintyProfit1921, especially Chapter 1 for a brief discussion closely related to the current subject].

Now, let a prospect be a *probability space* denoted by the tuple $(\Omega, \mathcal{F}, P)$. 
Following the @kolmogorovFoundationsTheoryProbability1950 axioms, $\Omega$ is the sample space

$$
\Omega = \{\omega_1, ..., \omega_n\}
\; ,
\tag{2.1}
$$

containing a finite set of possible outcomes $\omega$. 
$\mathcal{F}$ is the set of all possible subsets of $\Omega$, i.e., the power set $\mathcal{P}(\Omega)$.^[Note that this paper considers only the case of finite sample spaces containing of a small number of outcomes. The exact axiomatic definition of $\mathcal{F}$, which can be found in Kolmogorov [-@kolmogorovFoundationsTheoryProbability1950, pp. 2-3 and 14-15], also covers infinite sample spaces and is that of a $\sigma$-algebra. For the finite case, however, $\mathcal{P}(\Omega)$ can be considered as such. See Georgii [-@georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1.1] for a comprehensive but accessible mathematical treatment.]

$P$ is a probability measure

$$
P: \mathcal{F} \mapsto [0,1]
\; ,
\tag{2.2}
$$

which assigns each subset $A \in \mathcal{F}$, also referred to as *events*, a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.^[The probability measure formally assigns probabilities to sets of outcomes $A \in F$ rather than the elementary outcomes $\omega \in \Omega$. The probabilities of $\omega$ are determined by the probability mass function $p : \Omega \mapsto [0,1]$, which is used to construct the probability measure according to the equation $P(A) = \sum_{\omega \in A} p(\omega)$ for all $A \in \mathcal{F}$ [cf. @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1.2].]

In less technical terms, following the choice of a prospect, one of the outcomes associated with this particular prospect is obtained, a gain or loss respectively.
If there is just one outcome associated with a prospect, it is obtained with certainty.
However, if there is a collection of outcomes, just one outcome out of this collection is obtained according to some probability, where the probabilities can vary between outcomes and are given by the measure of the form described in (2.2) (see also Footnote 2).

Since it was stated that the inputs to the decision making process can be organized in a set of choice alternatives, we can adapt (1) for the case of a choice between $1, ..., k$ prospects to obtain:

$$
\iota : I' \mapsto I, \ \text{where} \ I = \{(\Omega_1,  \mathcal{F}_1, P_1), ..., (\Omega_k,  \mathcal{F}_k, P_k)\}
\; .
\tag{3}
$$

>>>>

The human mind is a cognitive system that operates on inputs from an environment or memory.
In the domain of decision making, such inputs provide information about the choice alternatives among which people can choose.
Let us treat these choice alternatives hereafter as elements $c$ of the set $C$.
To each $c \in C$ is then assigned some information from the inputs and this information may be used by the mind to learn about the *properties* of the respective choice alternative.
In turn, decision making can be considered an information processing problem which starts from the set $C$, containing all choice alternatives, and stops at a proper subset $C' \subset C$ with the inclusion map
$$
\begin{aligned}
  \iota : C' &\to C \\
  c &\mapsto c  \quad \forall c \in C'
\end{aligned}
\tag{1}
$$
[@jostMathematicalConcepts2015, p. 15], containing only the chosen alternatives.
Indeed, irrespective of their exact mathematical specification or computational implementation, decision models generally take as inputs information about a set of choice alternatives to return a subset thereof as outputs [@heOntologyDecisionModels2020].

@@

# Prospects, Decisions from Description, and Decisions from Experience

Since there is an uncountable number of decisions people confront in the natural world, behavioral decision research routinely abstracts from decisions between particular choice alternatives, e.g., the choice between job offers, political parties to vote for, investment plans, and whatnot.
Rather, with the choice between at least two *prospects* (see below), it studies a case which is counterfactual in that it omits the many particularities of each decision, but retains in the form of prospects the fundamental properties that almost all choice alternatives are assumed to possess [@lopesThoughtsPsychologicalConcept1983].^[See @knightRiskUncertaintyProfit1921, Chapter 1, for a timeless discussion of the reasonableness of such abstractions.] 
These fundamental properties are considered the possible outcomes of a choice alternative---hereafter $x_i$---and the probabilities with which these outcomes occur following the choice of the alternative---hereafter $p_X(x_i)$.
Roughly speaking, then, all choice alternatives that can be fully described by their outcome-probability pairs $(x_i, p_X(x_i))$ are prospects (or: gambles), which are exclusively considered below. 
<!--For illustration, take the choice alternatives in the following example with monetary outcomes:

@@

Now, more formally, let the axiomatic definition of a prospect be that of a discrete random variable's distribution.
Let therefore the tuple $(\Omega, \mathcal{F}, P)$ be a probability space. 
Following the @kolmogorovFoundationsTheoryProbability1950 axioms, $\Omega$ is the sample space containing a finite set of elements
$$
\omega_i \in \{\omega_1, ..., \omega_n\} = \Omega
\; .
\tag{2}
$$ 
$\mathcal{F}$ is the set of all possible subsets of $\Omega$, i.e., the power set $\mathcal{P}(\Omega)$ with each element $A \in P(\Omega) = \mathcal{F}$ being a subset of $\Omega$.^[Note that this paper considers only the case of finite sample spaces. The exact axiomatic definition of $\mathcal{F}$, which can be found in Kolmogorov [-@kolmogorovFoundationsTheoryProbability1950, pp. 2-3 and 14-15], also covers infinite sample spaces and is that of a $\sigma$-algebra. For the finite case, however, $\mathcal{P}(\Omega)$ can be considered as such [@georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1.1].] 
$P$ is a probability measure
$$
P: \mathcal{F} \to [0,1]
\; ,
\tag{3}
$$
which assigns each subset $A \in \mathcal{F}$ a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.<!--^[The probabilities $p(\omega)$ are determined by the probability mass function $p : \Omega \to [0,1]$, which is used to construct the probability measures $P(A) = \sum_{\omega \in A} p(\omega)$ for all $A \in \mathcal{F}$.]-->
The random variable $X$ is the measurable function
$$
\begin{aligned}
  X : (\Omega, \mathcal{F}) &\to (\Omega', \mathcal{F'})  \\
  \omega &\mapsto x
  \; ,
\end{aligned}
\tag{4}
$$
where $(\Omega', \mathcal{F'})$ is a measurable space with each element $x \in \Omega'$ being a possible outcome of a prospect and each element $A' \in \mathcal{F'}$ being a subset of $\Omega'$.
An outcome's probability $p_X(x)$ is then provided by the probability measure
$$
P_X : \mathcal{F'} \to [0,1]
\tag{5}
$$
with 
$$
P_X(A' \in \mathcal{F'}) := P(X^{-1} \ A') = P(\{\omega: X(\omega) = x \in A'\})
\tag{6}
$$
[see @kolmogorovFoundationsTheoryProbability1950, p. 21; see also @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 1, for a comprehensive introduction to random variables].
That is, the set consisting of a single outcome $x$ can be treated as a subset $\{x\} \in \mathcal{F'}$ and we therefore obtain $p_X(x) = P(X(\omega) = x)$.
Note that the probability measure in (5) and (6), respectively, is a discrete random variable's distribution and therefore a prospect.
To summarize and simplify, given a choice between $1, ..., k$ prospects, the inputs from the environment or memory should provide information about the set
$$
C = \{P_{X_1}, ... , P_{X_k}\}
\tag{7}
$$
with each element $c \in C$ being a distribution $P_X$ of a random variable defined on a different probability space.
Accordingly, each $c$ can itself be considered a set of outcome-probability pairs $(x_i, p_X(x_i)) \in \{(x_1, p_X(x_1)), ..., (x_n, p_X(x_n))\}$.

However, rarely in peoples' daily life, inputs take the form of explicit descriptions of all outcomes-probability pairs.
In such cases, people would make *decisions from description* (DfD), since the information from which the mind learns about the prospects are complete descriptions thereof.
Yet, rather often, people make *decisions from experience* [DfE, @hertwigDecisionsExperienceEffect2004], where the mind can learn about the prospects only from a record of past experiences with them.
That is, people may have experienced the possible outcomes of a prospect when they themselves or others have made the same decision repeatedly in the past, and the record of experienced outcomes for this particular prospect is a rough description of a *sample*.
The mind may then infer from the relative frequency with which an outcome occurred in such a sample of past decisions its probability of occurrence following a future decision.

Now, more formally, let the sample of a prospect be a set of realizations of the respective random variable $X$ according to its distribution $P_X$.^[Note that this paper considers the counterfactual case where samples are random, although it has been emphasized that in the natural world samples are "virtually never random" [@fiedlerBewareSamplesCognitiveecological2000, p. 660]. To account for non-random sampling, e.g., in the form of sampling biases [e.g., @fiedlerBewareSamplesCognitiveecological2000] or dynamic information structures [e.g., @cohenEffectPerceivedPatterns2021; @plonskyRelianceSmallSamples2015], the i.i.d-assumption would need to be dropped, which is out of the scope of this paper.]
So, for the choice between $1, ..., k$ prospects, the inputs to DfE are generated by a set $S$ of stochastic processes
$$
S = \{X_{1t}, ... , X_{kt}\}
\tag{8}
$$
with $t \in \{1, ..., N\}$ denoting a sequence of independent and identically distributed (i.i.d.) random variables and $N$ denoting the size of the respective sequence.
The sets of realizations that result from these stochastic processes may then be used to make an uncertain inference about the respective sets of outcome-probability pairs in (7).

Of course, the accuracy of such inferences depends, ceteris paribus, on the size of the samples, since from the laws of large numbers it follows that the relative frequency with which an outcome occurs in a sample converges to the outcome's objective probability only as the size of a sample increases [see @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, Chapter 5].
However, if the choices that people actually make between prospects are described, robust differences can be found between the choices in DfD and DfE, and the size of samples in DfE provides just one and only a partial explanation for this *description-experience gap* [see @hertwigDescriptionexperienceGapRisky2009; @wulffMetaanalyticReviewTwo2018]. 
The following section briefly addresses how choices between prospects in general and the description-experience gap in particular can be described using cumulative prospect theory's [CPT, @tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993] *weighting function*.^[This paper also considers CPT's value function, but emphasis is put on the weighting function.]
<!-- The two sections thereafter ... -->

# The Description of Choices in CPT and the Description-Experience Gap

The choices people actually make between prospects---be that in DfD or DfE---are often described in terms of deviations from expected value (EV) maximization, where the EV of a prospect is given by 
$$
EV = \sum_i^n x_i \times p_X(x_i)
\; .
\tag{9}
$$ 
That is, following the principle of EV maximization, one would calculate the prospects' EV---each by first multiplying its outcomes $x_i$ by their respective probabilities $p_X(x_i)$ and then summing the resulting products---to subsequently choose the prospect with the largest EV.
To *describe* how people's actual choices deviate from EV maximization [see @erevAnomaliesForecastsDescriptive2017, for a recent replication of classical demonstrations], the models of cumulative prospect theory [CPT, @tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993] and similar rank-dependent models [see @stottCumulativeProspectTheory2006, for an overview of models] fit choice data by maximizing the value
$$
V = \sum_i^n v(x_i) \times \pi_i
\; ,
\tag{10}
$$ 
with objective outcomes $x_i$ being transformed by a *value function* $v$, and *cumulative decision weights* $\pi_i$ being determined by the difference between transformed cumulative probabilities of the distribution $P_X$.^[Note that CPT or other originally deterministic decision models may be amended by a stochastic choice function to account for the probabilistic nature of choices [see, e.g., @rieskampProbabilisticNaturePreferential2008].]
Importantly, the transformation of cumulative probabilities is carried out by a weighting function $w$ (see (11) and (12) below).
More specifically, CPT proposes cumulative decision weights as replacements for objective probabilities to account for empirical regularities in the deviations from EV maximization in DfD^[@tverskyAdvancesProspectTheory1992 addressed the *fourfold pattern of risk attitudes*---i.e., choices between prospects indicate risk aversion for gains of high probability and losses of low probability, and risk seeking for gains of low probability and losses of high probability---and related effects such as the *certainty effect* and the *reflection effect* that were already addressed by CPT's predecessor, prospect theory [@kahnemanProspectTheoryAnalysis1979]. Together, these empirical choice patterns indicate a violation of the EV principle and the expected utility (EU) principle [see @bernoulliExpositionNewTheory1954\/1738].] that cannot be captured by the transformation of objective outcomes into subjective values $v(x)$ alone (see Footnote 6).
In essence, to provide a proper description of people's choices between prospects in terms of their deviation from EV maximization, the proposed model is such that the inferred cumulative decision weights may be larger or smaller than the respective objective probabilities in (9); and they may therefore give more or less weight to subjective values for the overall valuation of a prospect in CPT than would be objectively warranted.
In other words, for an inferred cumulative decision weight that is greater (smaller) than the objective probability it replaces, the respective outcome is overweighted (underweighted) in CPT.
Accordingly one may---but must not---adopt an as-if weighting terminology for the description of choices by stating that people decide *as if* they maximized the value in (10) and applied the inferred over- and underweighting pattern.
Importantly, the as-if prefix indicates that it cannot be concluded from the inferred cumulative decision weights that the mind does indeed perform the weighting computation in its literal sense, but it processes the inputs in a way that the outputs approximate the solutions of what can be considered probability weighting on an abstract computational level [see, e.g., @griffithsRationalUseCognitive2015].

Now, more formally, cumulative decision weights in CPT take the form
$$
\pi_i =
  \begin{cases}
     w^+(P(X \geq x_i)) - w^+(P(X > x_i)) \quad \forall x_i \geq 0, \\
     w^-(P(X \leq x_i)) - w^-(P(X < x_i)) \quad \text{else}, 
  \end{cases}
\tag{11}
$$
where $w$ is a monotonic, nonlinear function satisfying $w^+(0) = w^-(0) = 0$ and $w^+(1) = w^-(1) = 1$ [@tverskyAdvancesProspectTheory1992].
From the transformation of cumulative probabilities in (11), it follows that the replacement of objective probabilities by cumulative decision weights depends on the rank of the outcome to which the decision weight is assigned and the shape of the weighting function, which itself depends on the exact parameterization of the function [e.g., @goldsteinExpressionTheoryPreference1987; @tverskyAdvancesProspectTheory1992; @prelecProbabilityWeightingFunction1998; see also @stottCumulativeProspectTheory2006]. 
That is, the shape of the weighting function determines over which interval on the probability scale [0, 1] cumulative probabilities are overweighted or underweighted.
In turn, the same difference in the objective cumulative probability ...

Now, although the possible deviation patterns, which span the entire probability scale, depend on the exact parameterization of the cumulative weighting function, they allow the adoption of an as-if weighting terminology for the description of choice behavior, which was originally introduced outside of the CPT models [see @pachurWeightUncertainEvents2019].

Figure 1 illustrates such a description in CPT with different parameter settings for the two-parameter weighting function of @goldsteinExpressionTheoryPreference1987 with
$$
\begin{aligned}
  w : [0,1] &\to [0,1] \\
  p_X(x) &\mapsto \frac{\delta \times p_X(x)^{\gamma}}
  {\delta \times p_X(x)^{\gamma} + (1-p_X(x))^{\gamma}}
  \; ,
\end{aligned}
\tag{12}
$$
where $\gamma \in [0,2]$ and $\delta \geq 0$ determine the function's curvature and elevation, respectively.
In essence, if the estimate of a cumulative decision weight is larger (smaller) than the objective probability from which it is derived, the weighting function runs above (below) the dashed identity line.
The choice data are then described *as if* people maximized the value in (8) and applied the over- and underweighting pattern indicated by the estimated shape of the overall probability weighting function.

@@

ggplot(wf) +
  geom_line(aes(prob, w, group = gamma, color = gamma)) +
  geom_segment(x = 0, xend = 1, y =  0, yend = 1, linetype = "dashed", size = .2, color = "gray") +
  scale_color_scico(palette = "lisbon") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .2)) + 
  labs(x = "p",
       y = "w(p)", 
       color = expression(gamma)) + 
  facet_wrap(. ~ delta, labeller = label_bquote(cols = delta: .(delta))) + 
  theme_minimal()
```

@@

# Explanations for Underweighting in DfE

# Decision and Sampling Strategies

@@
